Landau quantization of nearly degenerate bands

### 2018-09-24 paper.tex
Updates from ShareLaTeX

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RD.png}
% alpha=0.05
% beta=0.02
% m=0.01
\caption{Rashba-Dresselhaus model. (a) Typical cyclotron orbits. (b) Comparison of Landau levels calculated by Eq. (\ref{eq:rule}) (black lines) and numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$ (red dots). Parameters for (b): $\alpha=7.5$($10^{5}$cm$\cdot \hbar$/s), $m=0.076m_e$, $\beta=3~(6)$($10^{5}$cm$\cdot \hbar$/s) for left (right) panels. (c) For various spin-orbit parameters $(\alpha,\beta)$ of the Rashba-Dresselhaus model and various field strength (satisfying $l^2S(E)\gg 1$), we plot  $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that lies closest to energy $E=0.4eV$, where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$. We also plot in (d) $l^2\delta S^2/S$, which has a strong positive relation with error in (c). Parameters for (c-d): $m=0.076m_e$. Units: $\alpha$($1.5\times 10^{5}$cm$\cdot \hbar$/s), $\beta$($1.5\times 10^{5}$cm$\cdot \hbar$/s), $Z$($10^{-3}$eV), $l^2 S$($10^2$).\label{fig:RD}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Black dashed lines illustrate a zeroth-order approximation of the Landau levels (assuming Klein tunneling); the leading-order correction is indicated by red lines. Exact Landau levels by numerical diagonalization are plotted as green dots. The Dirac point lies at $6.4$eV. Parameters for (b): $m$: 0.05$\hbar^2$/eV$\cdot$\AA$^2$, $\alpha$: 0.05eV$\cdot$\AA, $Z_\parallel$: 0.04eV. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\cite{RZfignote} Parameters for (d): $m=0.076m_e$, $l=100$\AA. \label{fig:RZ}}
\end{figure}

@@

As alluded to in \s{sec:qtznrules}, the adiabatic limit of two independent orbits does not exist if the two spin-orbit-split orbits exactly intersect at a II-Dirac point. To investigate this failure of adiabaticity, let us study the Landau levels at energies close to the II-Dirac point ($\epsilon_0$). We focus first on the regime where, on average (over $\frako_0$), the energy splitting $|\delta \var_+{-}\delta \var_-|$ [cf.\ \q{eq:RZ-Hamiltonian}] is much greater than $B_\perp|\calm_{\sma{+-}}|$. This implies that the split orbits do not appreciably hybridize -- except in the vicinity of the II-Dirac point where orbits (nearly) touch.  Near this touching point (where $t{=}0$), the dynamics of a spinor electron is described by linearizing $\H$ [cf.\ \q{calHRZ}] with respect to $t$:
\e{
\calh = -\alpha k_{E}\omega^{\sma{\perp}}_c t\,\sigma_x+(Z_\parallel-\alpha k_{E})\sigma_y-Z_\perp\sigma_z,\la{effhamIIDirac}
}
The eigenbasis of $\sigma_x$ corresponds to two diabatic levels with characteristic slope $v_d{:}{=}\alpha k_E\omega^{\sma{\perp}}_c$; near $t{=}0$, the diabatic levels anticross  due to the hybridization terms proportional to $\sigma_{y,z}$. Generally, the probability of tunneling across the hybridization-induced barrier is  $\rho^2{=}\exp(-2\pi\barmu)$, with $\barmu{=}{E_g}^2/{2v_d \hbar}$ and $E_g$ the energy gap at the center of the anticrossing;[Landau and lifshitz, Wittig] in this context,
\e{\barmu=\frac{Z_\perp^2+(Z_\parallel-\alpha k_{E})^2}{2\alpha k_{E}\var^{\sma{\perp}}_c}, \la{mu1}}
with $2\alpha k_E$ the spin-orbit splitting at energy $E$ [cf.\ \fig{fig:orbits}(a)]. In the absence of the out-of-plane Zeeman coupling ($Z_\perp{=}0$), $\bar{\mu}$ would vanish at the energy of the II-Dirac point -- leading to unit-probability Klein tunneling, as was first proposed by [Obrien] in a different model of the II-Dirac point. However, with a proper account of $Z_\perp$, we find instead that $\bar{\mu}$ has a nonvanishing minimum: $\bar{\mu}_{\sma{\text{min}}}{:}{=} ( m/m_0)(Z_\perp/4\alpha k_{E})$, which is a product of the dimensionless effective mass and the ratio of the out-of-plane Zeeman splitting over the spin-orbit splitting. This minimum is practically negligible for  semiconductor heterostructures with $m{\ll}m_0$, but may be relevant in heavy-fermion systems.

@@

For any nonzero field, $0{<}\tau{<}1$ implies that the two incommensurate harmonics $(\Omega_+{\pm}\Omega_-)$ in \q{coscos} compete to produce a quasirandom Landau-level spectrum [kaganovslutskin]. That is to say, the spectrum is disordered on the scale of nearest-neighbor level spacings but has longer-ranged correlations. To appreciate this, let us analyze the Landau levels (near the Dirac point) perturbatively in the small parameter $\tau$. To zeroth order, \q{coscos} reduces to
\e{2l^2S(E_n^0)+\pi = 2\pi n \rightarrow \{E_n^0(B)\}_{n\in\Z}.\label{trefoillerule}}
To derive the above equation, we have utilized $\phi^B_-{+}\phi^B_+{=}0$, which is valid for any two-band model in the quasidegenerate regime\footnote{For two band models, Berry phases of the two bands sum up to 0 on arbitrary path}; in \fig{fig:blochsphere} this is reflected in the sum of solid angles (at fixed energy) equalling $4\pi$. \q{trefoillerule} may be interpreted as a quantization rule for a trefoille-shaped orbit with net action $2S{=}4\pi m E{/\hbar^2}$; the corresponding Landau levels, as plotted with dashed lines in Fig.\ \ref{fig:RZ}(b), are  periodic in $E{\rightarrow}E{+}\pi/l^2(\partial S/\partial E)$ and in $l^2{\rightarrow}l^2{+}\pi/S$. Both periods are approximately half that of either disconnected spin-orbit-split orbit, in the assumed quasidegenerate regime $\delta S{\ll}S$. The $\pi$ correction in \q{trefoillerule} is the Berry phase associated to a $2\pi$ rotation of the wavefunction spin over a single trefoille orbit, as illustrated in Fig. \ref{fig:RZ}(a). The Maslov phase, which is generally equal to $\pi$ times the rotation number of an orbit, is trivially $2\pi$ for the trefoille orbit.

@@

To recapitulate, $E_n^0(B)$ [cf.\ \q{trefoillerule}] exhibits a periodicity on short scales associated to the large action of the trefoille orbit. This small-scale periodicity is lost due to  finite-$\tau$ corrections, yet a longer-range correlation (associated to the smaller differential action $\delta S$ between spin-split orbits) persists, as shown by the first-order-in-$\tau$ correction:\footnote{The general form of this perturbative calculation can be found in Sec. IX-E of [100page].} 
\e{&\delta E_n(B) = \f{(-1)^{n+1}}{2\pi} \tau \var_c \cos\bigg(\f{l^2\delta S}{{2}}+\f{\phi_+^B-\phi_-^B}{2}-\bar{\varphi}\bigg) \lin
    &\substack{\sma{Z_{\perp}{\rightarrow}0}\\ \approx} \f{(-1)^{n+1}}{4\sqrt{2\pi}}  \var_c \f{(E-\epsilon_0)l}{\alpha(\epsilon_0/Z_{\parallel})^{3/2}}\cos\bigg(\f{l^2\delta S}{{2}}-\bar{\varphi}\bigg),\label{RZfirstorderE}}
with the right-hand side (of each line) evaluated at $E_n^0(B)$. The second line of \q{RZfirstorderE} has greater utility where $m{\ll}m_0$. \red{This first order corrections of Landau levels are plotted in Fig. \ref{fig:RZ}(b) as red solid lines. The oscillation in Eq. (\ref{RZfirstorderE}) cannot be observed due to the smallness of $\delta S$. However, as expected from Eq. (\ref{RZfirstorderE}), first order correction have alternative signs for every Landau levels, which is confirmed by exact diagonalization [cf. Fig. \ref{fig:RZ}](b).}

If the quasidegenerate orbits are appreciably hybridized ( $|\delta \var_+{-}\delta \var_-|{\lesssim}B_\perp|\calm_{\sma{+-}}|$ on average), our quantization rule \qq{eq:rule}{eq:H1} remains valid but the propagator $\cala$ no longer has the closed analytic form of \q{propinplanezeeman}. We may anyway diagonalize $\cala$ numerically, and compare the resultant semiclassical Landau levels with the exact diagonalization of $H_{RZ}(\boldsymbol{K})$; to our knowledge, there exists no analytic solution of $H_{RZ}(\boldsymbol{K})$ for generic values of $\alpha,\zpar$ and $B_{\perp}$. For fixed  $B_{\perp}$ and a range of $(\alpha,\zpar)$ wherein  $l^2\delta S^2/S$ varies over $[0,0.2]$, the difference between semiclassical and exact Landau levels never exceeds  0.3\% of the cyclotron energy $\var_c$; this difference shows a strong positive correlation with $l^2\delta S^2/S$ [compare the left and right panels of Fig.\ \ref{fig:RZ} (d)], just like in the Rashba-Dresselhaus model.

For the special case of $\alpha{=}0$, the spectra of $H_{RZ}(\boldsymbol{K})$ and $\mathcal{A}[\mathfrak{o}]$ are both exactly soluble, hence it is possible to verify that the exact and semiclassical Landau levels match perfectly for every level index $n$ without any approximation; the latter is obtained from our rule [Eq.\ (\ref{eq:rule})], with the eigenphases of $\A$ equal to 
\e{\lambda_{\pm}{=}{\pm} 2\pi ml^2\sqrt{\f{Z_\parallel^2}{\hbar^4}{+}\f{g_\text{eff}^2}{16m^2l^4}}.}

@@

Suppose the propagator $\A$ were degenerate at $(\bar{E},\bar{B})$, i.e., $\lambda_1(\bar{E},\bar{B})=\lambda_2(\bar{E},\bar{B})$ mod $2\pi$. Generically, $(\bar{E},\bar{B})$ is not \emph{simultaneously} a solution of the quantization rule in \q{eq:rule}. That is to say, $\bar{E}$ is not generically the energy of a Landau level at field $\bar{B}$. To obtain a Landau level, we may tune either $E$ or $B$   away from $(\bar{E},\bar{B})$, and  utilize that  $\lambda_{\pm}$ are slowly-varying functions of $(E,B)$ compared to $l^2S(E)$:
\e{l^2\left|\p{S}{E}\right| \gg \left|\p{\lambda_a}{E}\right|,\, \left|S\right| \gg \left|\p{\lambda_a}{l^2}\right|;} 
these inequalities are guaranteed by the quasidegeneracy condition ($\delta S{\ll}S$) and the smoothness of $S(E)$, as shown in \app{sec:slowvariation}.  For fixed $\bar{B}$, we are therefore guaranteed to find a number (of order one) of paired Landau levels  in close proximity to $\bar{E}$, with each pair having an energetic splitting of order:
\e{ \bigg|\f{E_{1}-E_2}{\var_c}\bigg|_{B=\bar{B}} = \order\left( \f{\partial \lambda_a/\partial E}{\bar{l}^2\partial  S/\partial E}  \right)\ll 1.\la{llquasideg}}
Alternatively, we may fix $\bar{E}$ and investigate how Landau levels disperse as a function of $B$; the analog of \q{llquasideg}  is the existence of paired Landau levels in close proximity to $\bar{l}^2$, with a splitting of order:
\e{  \bigg|\f{l^2_{+}-l^2_-}{T_{{l^2}}}\bigg|_{E=\bar{E}} = O\left( \f{1}{S}\p{\lambda_a}{l^2} \right)\ll 1,\;\; T_{{l^2}}:=\f{2\pi}{S(\bar{E})}.\la{llquasidegB} }
$\partial \lambda_a/\partial l^2$ is estimated in \app{sec:slowvariation}, and
$T_{l^2}$ above is the period of quantum oscillations (in the absence of spin-orbit coupling). To recapitulate, for every exact degeneracy of $\cala$, there must exist, in close proximity, a number of paired Landau levels which are exactly degenerate, or nearly degenerate in the sense of \qq{llquasideg}{llquasidegB}. Each such pair of Landau levels will be called \textit{quasidegenerate}.

@@

In the absence of symmetry,  eigenvalue-degeneracies of the unitary propagator $\A$ [cf. \q{eq:prop}] are attainable by tuning three real parameters. Analogously, it is well known (as the non-crossing rule in quantum theory[von Neumann,wigner]) that eigenvalue-degeneracies of complex, finite-dimensional matrix Hamiltonians are also attainable by tuning three real parameters.
 To prove the non-crossing rule for two-by-two unitaries, we begin by decomposing any such matrix as $\A{=}e^{i\phi}{\cal \bar{A}}$, such that  ${\cal \bar{A}}{\in}\text{SU}(2)$ being equal to $\pm I$ is a necessary and sufficient condition for an eigenvalue-degeneracy of $\A$. In the canonical parametrization of $\text{SU}(2)$ over the three-sphere $S^3$: 
\e{{\cal \bar{A}}{=}\matrixtwo{r_1{+}ir_2}{r_3{+}ir_4}{{-}r_3{+}ir_4}{r_1{-}ir_2},\;\; \sum_{j=1}^4r_j^2{=}1,\;\; r_j\in \mathbb{R},  \la{s3}}
${\cal \bar{A}}{=}{\pm}I$ lie on distinct points of $S^3$, and are attainable by tuning three real parameters: $r_3{=}r_4{=}0$, and the phase of $r_1{+}ir_2$ must equal $0$ or $\pi$. The conclusion that three parameters need be tuned remains valid for any  non-singular parametrization of SU$(2)$.

@@

At finite field (assumed to be homogeneous in space and time), any remnant  symmetry (denoted $g$) must preserve the component of the magnetic moment in the direction of  $\bB$. The magnetic moment is odd under time reversal, and behaves as a pseudovector under spatial transformations.[sakurai] Therefore, either (i) $g$  preserves both the direction of time and the magnetic moment, or (ii) $g$ inverts both said objects. (i) is exemplified by a rotational symmetry with the rotational axis aligned parallel to the field, and (ii) by the composition  of time reversal $T$ and a spatial reflection $\mathfrak{r}_x$ (with reflection axis perpendicular to $\vec{B}$).

@@

In this and the subsequent section [\s{sec:rotsymmbreakdown}], we will demonstrate how rotational symmetry reduces codimension in the Landau-level problem. Before delving into this, it is instructive to consider the conceptually-simplest analog in differential geometry: if we parametrize the 2D plane by Cartesian coordinates $(x,y){\in}\R^2$, it is apparent that two parameters must be tuned to arrive at the origin. If however the origin is the center of a rotation symmetry, we may instead utilize polar coordinates and tune a single parameter (the radial coordinate) to arrive at $x{=}y{=}0$. We see that rotational symmetry reduces the codimension of the origin from two to one. Here, and in the  Landau-level problem, we define codimension with respect to a \textit{symmetrically-chosen coordinates} (if symmetry exists). We do not require that the symmetric coordinates is globally defined, e.g., the polar coordinates is not well-defined where the radial coordinate vanishes.

@@

The above discussion is reminiscent of topological (semi)metals which host Dirac-Weyl points and/or line nodes (of Bloch-Hamiltonian degeneracies). Weyl points in 3D $\bk$-space have codimension three and do not require any crystallographic point-group symmetry for their stability. Just like Weyl points, the point degeneracy of $\A$ is associated with a nontrivial Chern number (defined as an integral of the Berry curvature\footnote{The curvature is defined with respect to a line bundle formed from the nondegenerate eigenvector of $\cala$.} over a Gaussian surface surrounding the point). On the other hand,  both Dirac points (in 2D $\bk$-space, e.g., graphene) and  line nodes (in 3D $\bk$-space) have codimension two, hence their existence relies on point-group symmetries.

>>>>

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RD.png}
% alpha=0.05
% beta=0.02
% m=0.01
\caption{Rashba-Dresselhaus model. (a) Typical cyclotron orbits. (b) Comparison of Landau levels calculated by Eq. (\ref{eq:rule}) (black lines) and numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$ (red dots). Parameters for (b): $\alpha=7.5$($10^{5}$cm/s$\cdot\hbar$), $m=0.076m_e$, $\beta=3~(6)$($10^{5}$cm/s$\cdot\hbar$) for left (right) panels. (c) For various spin-orbit parameters $(\alpha,\beta)$ of the Rashba-Dresselhaus model and various field strength (satisfying $l^2S(E)\gg 1$), we plot  $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that lies closest to energy $E=0.4eV$, where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$. We also plot in (d) $l^2\delta S^2/S$, which has a strong positive relation with error in (c). Parameters for (c-d): $m=0.076m_e$. Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$), $\beta$($1.5\times 10^{5}$cm/s$\cdot\hbar$), $l^2 S$($10^2$).\label{fig:RD}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Oscillation of $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})] for even integer $n$ with respect to magnetic field calculated by Eq. (\ref{RZfirstorderE}) (black dashed line) and numerical diagonalization (red solid line). $\delta E$ is extracted at 0.45eV and the II-Dirac point is located at 0.5eV. Parameters for (b): $m$: 0.076$m_e$, $\alpha$: 1.5($10^{5}$cm/s$\cdot\hbar$), $Z_\parallel$: 1meV. Out of plane Zeeman coupling is ignored here. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\cite{RZfignote}, Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$). Parameters for (d): $m=0.076m_e$, $l=100$\AA. \label{fig:RZ}}.
\end{figure}

@@

As alluded to in \s{sec:qtznrules}, the adiabatic limit of two independent orbits does not exist if the two spin-orbit-split orbits exactly intersect at a II-Dirac point. To investigate this failure of adiabaticity, let us study the Landau levels at energies close to the II-Dirac point ($\epsilon_0$). We focus first on the regime where, on average (over $\frako_0$), the energy splitting $|\delta \var_+{-}\delta \var_-|$ [cf.\ \q{eq:RZ-Hamiltonian}] is much greater than $B_\perp|\calm_{\sma{+-}}|$. Equivalently, we are in the regime where $\alpha k_{\sma{E}}{\gg}\var_c$.\footnote{The orbit-averaged energy splitting is of order $\alpha k_{\sma{E}}$ for energies near the Dirac point. $B_\perp|\calm_{\sma{+-}}|$ is simply the Berry contribution to the Zeeman interaction, and has the form of the right-most term in \q{rashbaeffham}; this term is of order $\var_c$.} This implies that the split orbits do not appreciably hybridize -- except in the vicinity of the II-Dirac point where orbits (nearly) touch.  Near this touching point (where $t{=}0$), the dynamics of a spinor electron is described by linearizing $\H$ [cf.\ \q{calHRZ}] with respect to $t$:
\e{
\calh = -\alpha k_{E}\omega^{\sma{\perp}}_c t\,\sigma_x+(Z_\parallel-\alpha k_{E})\sigma_y-Z_\perp\sigma_z,\la{effhamIIDirac}
}
The eigenbasis of $\sigma_x$ corresponds to two diabatic levels with characteristic slope $v_d{:}{=}\alpha k_E\omega^{\sma{\perp}}_c$; near $t{=}0$, the diabatic levels anticross  due to the hybridization terms proportional to $\sigma_{y,z}$. Generally, the probability of tunneling across the hybridization-induced barrier is  $\rho^2{=}\exp(-2\pi\barmu)$, with $\barmu{=}{E_g}^2/{2v_d \hbar}$ and $E_g$ the energy gap at the center of the anticrossing;[Landau and lifshitz, Wittig] in this context,
\e{\barmu=\frac{Z_\perp^2+(Z_\parallel-\alpha k_{E})^2}{2\alpha k_{E}\var^{\sma{\perp}}_c}, \la{mu1}}
with $2\alpha k_E$ the spin-orbit splitting at energy $E$ [cf.\ \fig{fig:orbits}(a)]. In the absence of the out-of-plane Zeeman coupling ($Z_\perp{=}0$), $\bar{\mu}$ would vanish at the energy of the II-Dirac point -- leading to unit-probability Klein tunneling, as was first proposed by [Obrien] in a different model of the II-Dirac point. However, with a proper account of $Z_\perp$, we find instead that $\bar{\mu}$ has a nonvanishing minimum: $\bar{\mu}_{\sma{\text{min}}}{:}{=} ( m/m_0)(Z_\perp/4\alpha k_{E})$, which is a product of the dimensionless effective mass and the ratio of the out-of-plane Zeeman splitting over the spin-orbit splitting. This minimum is practically negligible for  semiconductor heterostructures with $m{\ll}m_0$, but may be relevant in heavy-fermion systems.

@@

For any nonzero field, $0{<}\tau{<}1$ implies that the two incommensurate harmonics $(\Omega_+{\pm}\Omega_-)$ in \q{coscos} compete to produce a quasirandom Landau-level spectrum [kaganovslutskin]. That is to say, the spectrum is disordered on the scale of nearest-neighbor level spacings but has longer-ranged correlations. To appreciate this, let us analyze the Landau levels (near the Dirac point) perturbatively in the small parameter $\tau$. To zeroth order, \q{coscos} reduces to
\e{2l^2S(E_n^0)+\pi = 2\pi n \rightarrow \{E_n^0(B)\}_{n\in\Z}.\label{trefoillerule}}
To derive the above equation, we have utilized $\phi^B_-{+}\phi^B_+{=}0$, which is valid for any two-band model in the quasidegenerate regime\footnote{For two band models, Berry phases of the two bands sum up to 0 on arbitrary path}; in \fig{fig:blochsphere} this is reflected in the sum of solid angles (at fixed energy) equalling $4\pi$. \q{trefoillerule} may be interpreted as a quantization rule for a trefoille-shaped orbit with net area $2S{=}4\pi m E{/\hbar^2}$; the corresponding Landau levels are periodic in $E{\rightarrow}E{+}\pi/l^2(\partial S/\partial E)$ and in $l^2{\rightarrow}l^2{+}\pi/S$. Both periods are approximately half that of either disconnected  orbit, in the assumed quasidegenerate regime $\delta S{\ll}S$. The $\pi$ correction in \q{trefoillerule} is the Berry phase associated to a $2\pi$ rotation of the wavefunction spin over a single trefoille orbit, as illustrated in Fig. \ref{fig:RZ}(a). The Maslov phase, which is generally equal to $\pi$ times the rotation number of an orbit, is trivially $2\pi$ for the trefoille orbit.

@@

To recapitulate, $E_n^0(B)$ [cf.\ \q{trefoillerule}] exhibits a periodicity on short scales associated to the large area of the trefoille orbit. This short-scale periodicity is lost due to  finite-$\tau$ corrections, yet a longer-range correlation (associated to the smaller differential action $\delta S$ between spin-split orbits) persists, as shown by the first-order-in-$\tau$ correction:\footnote{The general form of this perturbative calculation can be found in Sec. IX-E of [100page].} 
\e{&\delta E_n(B) = \f{(-1)^{n+1}}{2\pi} \tau \var_c \cos\bigg(\f{l^2\delta S}{{2}}+\f{\phi_+^B-\phi_-^B}{2}-\bar{\varphi}\bigg) \lin
    &\approx \f{(-1)^{n+1}}{4\sqrt{2\pi}}  \var_c \f{(E-\epsilon_0)l}{\alpha(\epsilon_0/Z_{\parallel})^{3/2}}\cos\bigg(\f{l^2\delta S}{{2}}-\bar{\varphi}\bigg),\label{RZfirstorderE}}
with the right-hand side (of each line) evaluated at $E{=}E_n^0(B)$. The second line is an approximation for $Z_{\perp}{\rightarrow}0$, and has greater utility where $m{\ll}m_0$. \red{Eq. (\ref{RZfirstorderE}) are compared with numerical diagonalization Fig. \ref{fig:RZ}(b) with even integers $n$.}

%( $|\delta \var_+{-}\delta \var_-|{\lesssim}B_\perp|\calm_{\sma{+-}}|$ on average)

If the quasidegenerate orbits are appreciably hybridized ($\alpha k_{\sma{E}}{\lesssim}\var_c$), our quantization rule \qq{eq:rule}{eq:H1} remains valid but the propagator $\cala$ no longer has the closed analytic form of \q{propinplanezeeman}. We may anyway diagonalize $\cala$ numerically, and compare the resultant semiclassical Landau levels with the exact diagonalization of $H_{RZ}(\boldsymbol{K})$; to our knowledge, there exists no analytic solution of $H_{RZ}(\boldsymbol{K})$ for generic values of $\alpha,\zpar$ and $B_{\perp}$. For fixed  $B_{\perp}$ and a range of $(\alpha,\zpar)$ wherein  $l^2\delta S^2/S$ varies over $[0,0.2]$, the difference between semiclassical and exact Landau levels never exceeds  0.3\% of the cyclotron energy $\var_c$; this difference shows a strong positive correlation with $l^2\delta S^2/S$ [compare the left and right panels of Fig.\ \ref{fig:RZ} (d)], just like in the Rashba-Dresselhaus model.

%For the special case of $\alpha{=}0$, the spectra of $H_{RZ}(\boldsymbol{K})$ and $\mathcal{A}[\mathfrak{o}]$ are both exactly soluble, hence it is possible to verify that the exact and semiclassical Landau levels match perfectly for every level index $n$ without any approximation; the latter is obtained from our rule [Eq.\ (\ref{eq:rule})], with the eigenphases of $\A$ equal to 
%\e{\lambda_{\pm}{=}{\pm} 2\pi ml^2\sqrt{\f{Z_\parallel^2}{\hbar^4}{+}\f{g_\text{eff}^2}{16m^2l^4}}.}

@@

Suppose the propagator $\A$ were degenerate at $(\bar{E},\bar{B})$, i.e., $\lambda_1(\bar{E},\bar{B})=\lambda_2(\bar{E},\bar{B})$ mod $2\pi$. Generically, $(\bar{E},\bar{B})$ is not \emph{simultaneously} a solution of the quantization rule in \q{eq:rule}. That is to say, $\bar{E}$ is not generically the energy of a Landau level at field $\bar{B}$. To obtain a Landau level, we may tune either $E$ or $B$   away from $(\bar{E},\bar{B})$, and  utilize that  $\lambda_{\pm}$ are slowly-varying functions of $(E,B)$ compared to $l^2S(E)$:
\e{l^2\left|\p{S}{E}\right| \gg \left|\p{\lambda_a}{E}\right|,\, \left|S\right| \gg \left|\p{\lambda_a}{l^2}\right|;} 
these inequalities are guaranteed by the quasidegeneracy condition ($\delta S{\ll}S$) and the smoothness of $S(E)$, as shown in \app{sec:slowvariation}.  For fixed $\bar{B}$, we are therefore guaranteed to find a number (of order one) of paired Landau levels  in close proximity to $\bar{E}$, with each pair having an energetic splitting of order:
\e{ \bigg|\f{E_{1}-E_2}{\var_c}\bigg|_{B=\bar{B}} = \order\left( \f{\partial \lambda_a/\partial E}{\bar{l}^2\partial  S/\partial E}  \right)\ll 1.\la{llquasideg}}
Alternatively, we may fix $\bar{E}$ and investigate how Landau levels disperse as a function of $B$; the analog of \q{llquasideg}  is the existence of paired Landau levels in close proximity to $\bar{l}^2$, with a splitting of order:
\e{  \bigg|\f{l^2_{+}-l^2_-}{T_{{l^2}}}\bigg|_{E=\bar{E}} = O\left( \f{1}{S}\p{\lambda_a}{l^2} \right)\ll 1,\;\; T_{{l^2}}:=\f{2\pi}{S(\bar{E})}.\la{llquasidegB} }
$\partial \lambda_a/\partial l^2$ is estimated in \app{sec:slowvariation}, and
$T_{l^2}$ above is the period (in $l^2$) of quantum oscillations -- in the absence of spin-orbit coupling. To recapitulate, for every exact degeneracy of $\cala$, there must exist, in close proximity, a number of paired Landau levels which are exactly degenerate, or nearly degenerate in the sense of \qq{llquasideg}{llquasidegB}. Each such pair of Landau levels will be called \textit{quasidegenerate}.

@@

In the absence of symmetry,  eigenvalue-degeneracies of the unitary propagator $\A$ [cf. \q{eq:prop}] are attainable by tuning three real parameters. Analogously, it is well known (as the non-crossing rule in quantum theory[von Neumann,wigner]) that eigenvalue-degeneracies of complex, finite-dimensional matrix Hamiltonians are also attainable by tuning three real parameters.
 To prove the non-crossing rule for two-by-two unitaries, we begin by decomposing any such matrix as $\A{=}e^{i\phi}{\cal \bar{A}}$, with  ${\cal \bar{A}}{\in}\text{SU}(2)$. Then the necessary and sufficient condition for an eigenvalue-degeneracy of $\A$ is that ${\cal \bar{A}}{=}{\pm}I$. In the canonical parametrization of $\text{SU}(2)$ over the three-sphere $S^3$: 
\e{{\cal \bar{A}}{=}\matrixtwo{r_1{+}ir_2}{r_3{+}ir_4}{{-}r_3{+}ir_4}{r_1{-}ir_2},\;\; \sum_{j=1}^4r_j^2{=}1,\;\; r_j\in \mathbb{R},  \la{s3}}
${\cal \bar{A}}{=}{\pm}I$ lie on distinct points of $S^3$, and are attainable by tuning three real parameters: $r_2{=}r_3{=}r_4{=}0$. The conclusion that three parameters need be tuned remains valid for any  non-singular parametrization of SU$(2)$.

@@

Any symmetry $g$ that remains at finite field (assumed to be homogeneous in space and time) must preserve the component of the magnetic moment in the direction of  $\vec{B}$. The magnetic moment is odd under time reversal, and behaves as a pseudovector under spatial transformations.[sakurai] Therefore, either (i) $g$  preserves both the direction of time and the magnetic moment, or (ii) $g$ inverts both said objects. (i) is exemplified by a rotational symmetry with the rotational axis aligned parallel to the field, and (ii) by the composition  of time reversal $T$ and a spatial reflection $\mathfrak{r}_x$ (with reflection axis perpendicular to $\vec{B}$).

@@

In this and the subsequent section [\s{sec:rotsymmbreakdown}], we will demonstrate how rotational symmetry reduces codimension in the Landau-level problem. Before delving into this, it is instructive to consider the conceptually-simplest analog in differential geometry: if we parametrize the 2D plane by Cartesian coordinates $(x,y){\in}\R^2$, it is apparent that two parameters must be tuned to arrive at the origin. If however the origin is the center of a rotation symmetry, we may instead utilize polar coordinates and tune a single parameter (the radial coordinate) to arrive at $x{=}y{=}0$. We see that rotational symmetry reduces the codimension of the origin from two to one. Here, and in the  Landau-level problem, we define codimension with respect to a \textit{symmetrically-chosen coordinates} (if symmetry exists); such symmetric coordinates need not be well-defined at isolated high-symmetry points, just as the polar coordinates are not well-defined at the rotational center.

@@

%The above discussion is reminiscent of topological (semi)metals which host Dirac-Weyl points and/or line nodes (of Bloch-Hamiltonian degeneracies). Weyl points in 3D $\bk$-space have codimension three and do not require any crystallographic point-group symmetry for their stability. Just like Weyl points, the point degeneracy of $\A$ is associated with a nontrivial Chern number (defined as an integral of the Berry curvature\footnote{The curvature is defined with respect to a line bundle formed from the nondegenerate eigenvector of $\cala$.} over a Gaussian surface surrounding the point). On the other hand,  both Dirac points (in 2D $\bk$-space, e.g., graphene) and  line nodes (in 3D $\bk$-space) have codimension two, hence their existence relies on point-group symmetries.
